---
title: "IMDB Ratings: Steven Spielberg vs Tim Burton"
author: "Hugo Huere Alvarado"
date: '2020-10-11'
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)

```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
# Loading all libraries
library(tidyverse)
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
library(ggplot2)
library(patchwork)
library(infer)
library(vroom)
```

This project was done to research whether the IMDB rating for blockbuster movies were influenced by the director. Before looking at any kind of data, a null hypothesis and alternative hypothesis were set up as follows: 

### Null Hypothesis
There is no difference in the mean IMDB-rating for movies produced by Steven Spielberg or Tim Burton. 

### Alternative Hypothesis
There is a significant difference between these mean IMDB-ratings.

We will look at a subset sample of movies, taken from the [Kaggle IMDB 5000 movie dataset](https://www.kaggle.com/carolzhangdc/imdb-5000-movie-dataset)

```{r,load_movies, warning=FALSE, message=FALSE}
movies <- vroom::vroom(here::here("data", "movies.csv"))
glimpse(movies)
```
Besides the obvious variables of `title`, `genre`, `director`, `year`, and `duration`, the rest of the variables are as follows:

- `gross` : The gross earnings in the US box office, not adjusted for inflation
- `budget`: The movie's budget 
- `cast_facebook_likes`: the number of facebook likes cast memebrs received
- `votes`: the number of people who voted for (or rated) the movie in IMDB 
- `reviews`: the number of reviews for that movie
- `rating`: IMDB average rating 

Are there any missing values (NAs)? Are all entries distinct or are there duplicate entries?

**No missing but many duplicates across categories. There are 2961 rows and only 2907 unique movie titles, so it can be inferred that some movies show more than 1 times with different genres.**


Then I manipulate the data set so that only Steven Spielburg and Tim Burton are evaluated.

```{r}
movies_v1 <- movies %>% filter(director == "Steven Spielberg" | director == "Tim Burton") %>% select(director | rating)

movies_v2 <- movies_v1 %>%
  group_by(director) %>%
  summarise(mean_rating = mean(rating, na.rm=TRUE),
            sd_rating = sd(rating, na.rm = TRUE),
            count = n(),
            se_rating = sd_rating/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_rating,
            lower = mean_rating - t_critical * se_rating,
            upper = mean_rating + t_critical * se_rating
            )
```
Steven Spielburg has an average IMDB rating of 7.57 while Tim Burton has an average rating of just 6.93. Furthermore, Steven has made more movies with a smaller spread across ratings, which is seen by his standard error of just 0.148. On the other hand, Tim Burton has made fewer movies with a larger discrepancy in ratings; causing him to have an standard error of 0.193.

### Data Visualisation
Judging from these numbers alone, it does indeed seem that a movie's director has an impact on the IMDB Rating. But looking at numbers can be a bit daunting, so let's try to make this a bit more appealing to look at: 

``` {r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=13}
p <- ggplot(movies_v2, aes(mean_rating, reorder(director, mean_rating), colour = director)) +
  labs(title = "Do Spielberg and Burton have the same mean IMDB ratings?", subtitle = "95% confidence intervals overlap", x= "Mean IMDB Rating", y = "") +
  theme_bw() + theme(legend.position = "none") +
  geom_point(size = 7) +
  geom_text(aes(label = round(mean_rating, 2)), vjust = -1.2, size = 7, colour = "black") +
  geom_text(aes(x = upper, label = round(upper, 2)), vjust = -1.6, size = 5, colour = "black") +
  geom_text(aes(x = lower, label = round(lower, 2)), vjust = -1.6, size = 5, colour = "black") +
  geom_errorbar(movies_v2, mapping = aes(x=mean_rating, xmin=upper, xmax=lower, y = director), width = 0.1,    size = 2.5) +
  geom_rect(aes(xmin = lower[1], xmax = upper[2], ymin = 0, ymax = 3), colour = "grey", alpha = 0.2)

p

```


## Hypothesis test with formula

```{r}
movies <- movies %>%
  mutate(spielberg_burton = movies$director %in% c("Steven Spielberg", "Tim Burton"))

movies %>% filter(!is.na(spielberg_burton)) %>% 
  group_by(spielberg_burton) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count))

t.test(rating ~ spielberg_burton, data = movies)
```


## Hypothesis test with infer

```{r}

obs_diff <- movies %>%
  specify(rating ~ spielberg_burton) %>%
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
obs_diff


null_dist <- movies %>%
  specify(rating ~ spielberg_burton) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
null_dist

ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()
  
null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```

##  Conclusion
Using Michael Bay and James Cameron this becomes a lot more apparent - here the gap between the two confidence intervals is quite large and so **tells us that a movie's director most likely plays a role in its IMDB rating**.