```
title = "STEVEN SPIELBERG VS TIM BURTON?"
author = "Hugo Huere Alvarado"
date = "`r Sys.Date()`"
output =
  html_document:
    theme = flatly
    highlight = zenburn
    number_sections = yes
    toc = yes
    toc_float = yes
    code_folding = show
```

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
# Loading all libraries
library(tidyverse)
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
library(ggplot2)
library(patchwork)
library(infer)
```


# IMDB ratings: Differences between directors
This project was done to research whether the IMDB rating for blockbuster movies were influenced by the director. Before looking at any kind of data, a null hypothesis and alternative hypothesis were set up as follows: 

### Null Hypothesis
* There is no difference in the mean IMDB-rating for movies produced by Steven Spielberg or Tim Burton. 

### Alternative Hypothesis
* There is a significant difference between these mean IMDB-ratings.

First, let us load the data into R:

```{r load-movies-data}
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)
```

# Your R code and analysis should go here. If you want to insert a blank chunk of R code you can just hit `Ctrl/Cmd+Alt+I` 
```{r}

## Reproducing the graph

movies_v1 <- movies %>% filter(director == "Steven Spielberg" | director == "Tim Burton") %>% select(director | rating)

movies_v2 <- movies_v1 %>%
  group_by(director) %>%
  summarise(mean_rating = mean(rating, na.rm=TRUE),
            sd_rating = sd(rating, na.rm = TRUE),
            count = n(),
            se_rating = sd_rating/sqrt(count),
            t_critical = qt(0.975, count-1), 
            margin_of_error = t_critical * se_rating,
            lower = mean_rating - t_critical * se_rating,
            upper = mean_rating + t_critical * se_rating
            )

p <- ggplot(movies_v2, aes(mean_rating, reorder(director, mean_rating), colour = director)) +
  labs(title = "Do Spielberg and Burton have the same mean IMDB ratings?", subtitle = "95% confidence intervals overlap", x= "Mean IMDB Rating", y = "") +
  theme_bw() + theme(legend.position = "none") +
  geom_point(size = 7) +
  geom_text(aes(label = round(mean_rating, 2)), vjust = -1.2, size = 7, colour = "black") +
  geom_text(aes(x = upper, label = round(upper, 2)), vjust = -1.6, size = 5, colour = "black") +
  geom_text(aes(x = lower, label = round(lower, 2)), vjust = -1.6, size = 5, colour = "black") +
  geom_errorbar(movies_v2, mapping = aes(x=mean_rating, xmin=upper, xmax=lower, y = director), width = 0.1,    size = 2.5) +
  geom_rect(aes(xmin = lower[1], xmax = upper[2], ymin = 0, ymax = 3), colour = "grey", alpha = 0.2)

p


## Hypothesis test with formula

movies <- movies %>%
  mutate(spielberg_burton = movies$director %in% c("Steven Spielberg", "Tim Burton"))

movies %>% filter(!is.na(spielberg_burton)) %>% 
  group_by(spielberg_burton) %>% 
  summarise(count = n()) %>% 
  mutate(prop= count/sum(count))

t.test(rating ~ spielberg_burton, data = movies)

## Hypothesis test with infer

obs_diff <- movies %>%
  specify(rating ~ spielberg_burton) %>%
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
obs_diff


null_dist <- movies %>%
  specify(rating ~ spielberg_burton) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c(TRUE, FALSE))
null_dist

ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()
  
null_dist %>% visualize() +
  shade_p_value(obs_stat = obs_diff, direction = "two-sided")

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")
```
